---
title: "Deliverable 1"
author: "Pere Arnau Alegre & Andrés Jiménez González"
date: \today
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 4
  word_document:
    toc: no
    toc_depth: '4'
  html_document:
    toc: no
    toc_depth: '4'
geometry: left=1.9cm,right=1.9cm,top=1.25cm,bottom=1.52cm
fontsize: 18pt
subtitle: Data Processing, Description, Validation and Profiling
classoption: a4paper
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Data description
* Description https://www.kaggle.com/datasets/adityadesai13/used-car-dataset-ford-and-mercedes
* Data Dictionary - Scraped data of used cars, which have been separated
into files corresponding to each car manufacturer (only Mercedes, BMW,
Volkswagen and Audi cars are to be considered).

## Variables
* Model
  * A string indicating the model of the car.     
* Year	
  * A discrete numeric variable to indicate the year the car was sold
* Price
  * Continuous variable indicating the price at which the car was sold
* Transmission
  * Categorical variable that indicates the type of transmission of the car
  * Values:
    * Automatic
    * Manual
    * Semi-Automatic
    * Other
* Mileage
  * A discrete numeric variable to indicate the number of miles the car had when it was sold
* Fuel Type
  * Categorical variable that indicates the type of fuel of the car
  * Values:
    * Diesel
    * Electric
    * Hybrid
    * Petrol
    * Other
* Tax
  * A discrete numeric variable to indicate the road tax of the vehicle.
* MPG
  *  Continuous variable indicating the fuel consumption of the car
* Engine Size
  * Continuous variable indicating the size of the engine
* Manufacturer
  * Categorical variable that indicates the manufacturer brand of the car.
  * Values:
    * Mercedes
    * Audi
    * Volkswagen
    * BMW


# Loading of Required Packages for the deliverable
We load the necessary packages and set the working directory
```{r echo = T, results = 'hide', message=FALSE, error=FALSE, warning=FALSE}
#setwd("C:/Users/TOREROS-II/Documents/ANDRES/UNI/ADEI/trabajo/deliverable1")
setwd("C:/Users/Arnau/Desktop")
# Load Required Packages
options(contrasts=c("contr.treatment","contr.treatment"))
requiredPackages <- c("missMDA","chemometrics","mvoutlier","effects","FactoMineR","car", "factoextra","RColorBrewer","dplyr","ggmap","ggthemes","knitr", "corrplot")
missingPackages <- requiredPackages[!(requiredPackages %in% installed.packages()[,"Package"])]
if(length(missingPackages)) install.packages(missingPackages)
lapply(requiredPackages, require, character.only = TRUE)
```  

## Select a sample of 5000 records
From the proposed database, we need to select a sample of 5000 records randomly so we can start analyzing our data.
```{r echo = T, results = 'hide'}
if(!is.null(dev.list())) dev.off()  # Clear plots
rm(list=ls())                       # Clean workspace
```


Data: used_car_dataset.csv
```{r}
#filepath<-"C:/Users/TOREROS-II/Documents/ANDRES/UNI/ADEI/trabajo/deliverable1"
filepath<-"C:/Users/Arnau/Desktop/"
df<-read.table(paste0(filepath,"/sample_5000.csv"),header=T, sep=",")[c(-1)]

# dim(df)       # Displays the sample size
# names(df)     # Displays the names of the sample variables
# summary(df)   
```


## Some useful functions
```{r}
calcQ <- function(x) { # Function to calculate the different quartiles
  s.x <- summary(x)
  iqr<-s.x[5]-s.x[2]
  list(souti=s.x[2]-3*iqr, mouti=s.x[2]-1.5*iqr, min=s.x[1], q1=s.x[2], q2=s.x[3], 
       q3=s.x[5], max=s.x[6], mouts=s.x[5]+1.5*iqr, souts=s.x[5]+3*iqr ) 
}
countNA <- function(x) { # Function to count the NA values
  mis_x <- NULL
  for (j in 1:ncol(x)) {mis_x[j] <- sum(is.na(x[,j])) }
  mis_x <- as.data.frame(mis_x)
  rownames(mis_x) <- names(x)
  mis_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {mis_i <- mis_i + as.numeric(is.na(x[,j])) }
  list(mis_col=mis_x,mis_ind=mis_i) 
}
countX <- function(x,X) { # Function to count a specific number of appearences
  n_x <- NULL
  for (j in 1:ncol(x)) {n_x[j] <- sum(x[,j]==X) }
  n_x <- as.data.frame(n_x)
  rownames(n_x) <- names(x)
  nx_i <- rep(0,nrow(x))
  for (j in 1:ncol(x)) {nx_i <- nx_i + as.numeric(x[,j]==X) }
  list(nx_col=n_x,nx_ind=nx_i) 
}
```

#Univariate Description and Preprocessing


## Qualitative Variables (Factors) / Categorical
**Description**: Original numeric variables corresponding to qualitative concepts have to be converted to factors. New factors grouping original levels will be considered very positively.
We need to do an analysis of all the variables to be able to identify missings, errors and outliers. 
We will also try to factorize each variable to make it easier to understand the sample.

# Variable initialization of missings, outliers and errors for columns
```{r}
jmis<-rep(0,2*ncol(df))  # columns - variables

mis1<-countNA(df)
#mis1$mis_ind   # Number of missings for the current set of cars (observations)
#mis1$mis_col   # Number of missings for the current set of variables

jouts<-rep(0,ncol(df))  # columns - variables

jerrs<-rep(0,ncol(df))  # columns - variables
```

### 3. Price
We know that the price should be positive, so we will treat as errors the prices <= 0. We don't count the errors by rows for the variable price because we erase that rows.
```{r}
summary(df$price)
sel<-which(df$price <= 0)
jerrs[which(colnames(df)=="price")] <- length(sel)
#We will delete the rows with errors in the price because we cannot make imputations for our target variable.
df <- df[which(df$price > 0), ]

boxplot(df$price)
var_out<-calcQ(df$price)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")
#We can see there are outliers in the dataset so we will treat them.
#As this is the response variable, we will delete the outlier rows because we cannot delete the value and impute it.
llout_price<-which((df$price > var_out$souts) | (df$price < var_out$souti ))
#iouts[llout] <- iouts[llout]+1
jouts[which(colnames(df)=="price")]<-length(llout_price)
df <- df[-llout_price, ]
```

# Variable initialization of missings, outliers and errors for rows
Initialization of counts for missings, outliers and errors. All numerical variables have to be checked before.
```{r}
imis<-rep(0,nrow(df))  # rows - cars

iouts<-rep(0,nrow(df))  # rows - cars

ierrs<-rep(0,nrow(df))  # rows - cars
```

### 1. Model
This variable indicates the model of the car.
```{r}
df$model<-factor(paste0(df$manufacturer,"-",df$model))
#levels(df$model)
summary(df$model)
#Too many models to represent them in a graph
#The is not missing data or erroneous data, so we will not make any change in the model column
```

### 2. Year
A discrete numeric variable to indicate the year the car was sold, ranging from 1970 to 2020

#### Factorization and outlier detection of the variable year
```{r}
boxplot(df$year, main="Boxplot of sold year", col="darkslateblue")
#df$year <- factor(df$year)
#We can see that there are outliers in the dataset, so we will treat them.
summary(df$year)
var_out<-calcQ(df$year)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

llout<-which((df$year <= var_out$souti))
iouts[llout] <- iouts[llout]+1
jouts[which(colnames(df)=="year")]<-length(llout)

#We will group all the inferior outliers into one variable
df[llout,"year"] <- NA
summary(df$year)
#df[which(df$year<=var_out$souti),"year"] <- paste0(var_out$souti, " or before")
barplot(table(df$year), main="Barplot of sold year", col="darkslateblue")
```

In orther to better analyze the price of the cars and to group them, we will create a categorical variable representing the price of the car.
```{r}
df$price_type <- df$price
df$price_type[which(df$price >= var_out$min & df$price_type < var_out$q1)] <- "super cheap"
df$price_type[which(df$price >= var_out$q1 & df$price_type < var_out$q2)] <- "cheap"
df$price_type[which(df$price >= var_out$q2 & df$price_type < var_out$q3)] <- "expensive"
df$price_type[which(df$price >= var_out$q3 & df$price_type < var_out$mouts)] <- "very expensive"
df$price_type[which(df$price >= var_out$mouts )] <- "extremely expensive"
table(df$price_type)
```


### 4. Transmission

```{r}
df$transmission <- factor( df$transmission )
levels( df$transmission )
df$transmission <- factor( df$transmission, levels = c("Manual","Semi-Auto","Automatic"),labels = paste0("f.Trans-",c("Manual","SemiAuto","Automatic")))
#All transmission not listed above have been replaced as NA
```

### 5. Mileage
```{r}
boxplot(df$mileage)
var_out<-calcQ(df$mileage)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")
llout_mil<-which((df$mileage<var_out$souti)|(df$mileage>var_out$souts))
iouts[llout_mil]<-iouts[llout_mil]+1
df[llout_mil,"mileage"] <- NA
```


### 6. fuelType
Andres
```{r}
df$fuelType <- factor(df$fuelType)
levels(df$fuelType)
df$fuelType <- factor( df$fuelType, levels = c("Diesel","Petrol","Hybrid"), labels = paste0("f.Fuel-",c("Diesel","Petrol","Hybrid")))
#All fuelTypes not listed above have been replaced as NA
```


### 7. Tax
Andres
```{r}
boxplot(df$tax, main="Boxplot of tax", col="darkslateblue")
#df$year <- factor(df$year)
#We can see that there are outliers in the dataset, so we will treat them.
summary(df$tax)
var_out<-calcQ(df$tax)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")

llout<-which((df$tax <= var_out$souti & df$tax >= var_out$souts))
iouts[llout] <- iouts[llout]+1
jouts[which(colnames(df)=="tax")]<-length(llout)
df[llout, "tax"] <- NA

summary(df$tax)
boxplot(df$tax)

```

### 8. MPG
Andres
```{r}
#Outliers are replaced by NA
boxplot(df$mpg)
var_out<-calcQ(df$mpg)
abline(h=var_out$souts,col="red")
abline(h=var_out$souti,col="red")
llout_mpg<-which((df$mpg<var_out$souti)|(df$mpg>var_out$souts))
iouts[llout_mpg]<-iouts[llout_mpg]+1
jouts[which(colnames(df)=="mpg")]<-length(llout)
df[llout_mpg,"mpg"] <- NA
```


### 9. EngineSyze
Andres --> contabilizar errores
```{r}
df$engineSize <- factor(df$engineSize)
levels(df$engineSize)

df[which(df[,"engineSize"]==0),]

# It is a quantitive variable  Non-possible values will be recoded to NA
sel<-which(df$engineSize==0)
ierrs[sel]<-ierrs[sel]+1 #Vector of errors per individual update
sel                 #### sel contains the rownames of the individuals with "0" 
#                        as  value for engineSize
# We should update jerrs vector: errors per variable

# df[sel,"engineSize"]<-3    # non-possible values are replaced by NA, missing value symbol in R
# NA assignment for forward imputation:
df[sel,"engineSize"]<-NA
```

#Imputation
What we do with imputation is be able to eliminate all those values that may be missings, outliers or errors to turn them into values that can be realistic within our sample.

## Imputation of numeric variables

```{r}
library(missMDA)
# Now one by one describe vars and put them on lists
vars_con<-c("year", "mileage", "tax", "mpg")
vars_res<-c("price")

summary(df[,vars_con])
res.impca<-imputePCA(df[,vars_con],ncp=3)
summary(res.impca$completeObs)

# Check one by one:
par(mfrow=c(1,2))
hist(df$year, main="Hist of year before imputation")
hist(res.impca$completeObs[,"year"], main="Hist of year after imputation")
hist(df$mileage, main="Hist of mileage before imputation")
hist(res.impca$completeObs[,"mileage"], main="Hist of mileage after imputation")
hist(df$tax, main="Hist of tax before imputation")
hist(res.impca$completeObs[,"tax"], main="Hist of tax after imputation")
hist(df$mpg, main="Hist of mpg before imputation")
hist(res.impca$completeObs[,"mpg"], main="Hist of mpg after imputation")

# Once you have validated the process:
df[,vars_con ]<-res.impca$completeObs

```

## Imputation of qualitative variables

```{r}
vars_dis<-c("model","transmission","fuelType","engineSize","manufacturer")
summary(df[,vars_dis])
res.immca<-imputeMCA(df[,vars_dis],ncp=4)
summary(res.immca$completeObs)
# Check one by one (we only have enginesize, transmission & fuelType)
barplot(table(df$engineSize), main="Barplot of engineSize before imputation")
barplot(table(res.immca$completeObs[,"engineSize"]), main="Barplot of engineSize after imputation")
barplot(table(df$transmission), main="Barplot of transmission before imputation")
barplot(table(res.immca$completeObs[,"transmission"]), main="Barplot of transmission after imputation")
barplot(table(df$fuelType), main="Barplot of fuelType before imputation")
barplot(table(res.immca$completeObs[,"fuelType"]), main="Barplot of fuelType after imputation")


# Once you have validated the process
df[ , vars_dis ]<-res.immca$completeObs

# Are there NA?
sum(countNA(df)$mis_ind)==0

par(mfrow=c(1,1))
```


# Creation and discretization of new variables

## 1. New variable: Audi/Not Audi
```{r}
# Binary Target: Audi?

df$Audi<-ifelse(df$manufacturer == "Audi",1,0)
df$Audi<-factor(df$Audi,labels=c("No","Yes"))
summary(df$Audi)
# Pie
piepercent<-round(100*(table(df$Audi)/nrow(df)),dig=2); piepercent
pie(table(df$Audi),col=terrain.colors(2),labels=paste(piepercent,"%"), main="Piechart of Audi cars")
legend("topright", levels(df$Audi), cex = 0.8, fill = terrain.colors(2))
# Bar Chart
barplot(table(df$Audi),main="Barplot Binary Outcome - Factor",col=terrain.colors(2))
```

## 2. New variable: yearsAferSell
A discrete numeric variable to indicate how many years have passed from when the car was sold since 2022.

```{r}
df$years_after_sell <-  2022 - df$year
boxplot(df$years_after_sell, main="Boxplot of years after sell", col="cyan3", ylim=c(0,15))
#There are no extreme outliers in the variable because we treated outliers in the variable year.
summary(df$years_after_sell)
```

## 3. Discretization of the variable Tax
```{r}
quantile(df$tax,seq(0,1,0.25),na.rm=TRUE)
quantile(df$tax,seq(0,1,0.1),na.rm=TRUE)
quants <- calcQ(df$tax)

# df$aux<-factor(cut(df$tax,breaks=quantile(df$tax,seq(0,1,0.25),na.rm=TRUE),include.lowest = T )) # Does not work
#Reconsiderations of limits bc mean and 3rd quantile are the same
#aux<-factor(cut(df$tax,breaks=c(0, 125, 145, quants),include.lowest = T ))
#summary(aux)
#tapply(df$tax,aux,median)
df$f.tax<-factor(cut(df$tax, breaks=c(quants$min,quants$q1, quants$q2, quants$q3+10, quants$max), include.lowest=T))
levels(df$f.tax)<-paste("f.tax-",levels(df$f.tax),sep="")
table(df$f.tax,useNA="always")
```

## 4. Discretization of the variable mileage
```{r}
df$f.mileage<-factor(cut(df$mileage,breaks=c(quantile(df$mileage,seq(0,1,0.25),na.rm=TRUE)),include.lowest = T ))
levels(df$f.mileage)<-paste("f.mileage-",levels(df$f.mileage),sep="")
table(df$f.mileage,useNA="always")
```

## 4. Discretization of the variable mpg
```{r}
df$f.mpg<-factor(cut(df$mpg,breaks=c(quantile(df$mpg,seq(0,1,0.25),na.rm=TRUE)),include.lowest = T ))
levels(df$f.mpg)<-paste("f.mpg-",levels(df$f.mpg),sep="")
table(df$f.mpg,useNA="always")
```

## 5. Discretization of the variable year
```{r}
df$f.year<-factor(cut(df$year,breaks=c(quantile(df$year,seq(0,1,0.25),na.rm=TRUE)),include.lowest = T ))
levels(df$f.year)<-paste("f.mpg-",levels(df$f.year),sep="")
table(df$f.year,useNA="always")
```

#Create variable adding the total number missing values, outliers and errors.
Describe these variables, to which other variables exist higher associations.

##Compute the correlation with all other variables. Rank these variables according the correlation
```{r}
df$inconsistencies <- imis+iouts+ierrs
vars_quanti <- c(2,3,5,7,8,13,18)
res <- cor(df[,vars_quanti])
round(res, 2)
corrplot(res)
```

##Mean of missing/outliers/errors per groups
Compute for every group of individuals (group of age, etc, …) the mean of missing/outliers/errors values. Rank the groups according the computed mean.
```{r}
dfInconsistencias <- data.frame("Modelo"=c("Audi", "VM", "Mercedes", "BMW"))
vw_inconsis <- mean(df$inconsistencies[ which(df$manufacturer=="VW")])
audi_inconsis <- mean(df$inconsistencies[ which(df$manufacturer=="Audi")])
bmw_inconsis <- mean(df$inconsistencies[ which(df$manufacturer=="BMW")])
merc_inconsis <- mean(df$inconsistencies[ which(df$manufacturer=="Mercedes")])
dfInconsistencias$incons <- c(vw_inconsis, audi_inconsis, bmw_inconsis, merc_inconsis)
dfInconsistencias<-dfInconsistencias[order(dfInconsistencias$incons),]
barplot(dfInconsistencias$incons, names.arg = dfInconsistencias$Modelo)

```

#Multivariant outliers
We don't use the variable tax for the searching of multivariant outliers because 
it is a column linearly dependent with other column.
```{r}
res.mout <- Moutlier( df[ ,c(2,3,5,8)], quantile = 0.995) 
par(mfrow = c(1,1))
plot( res.mout$md, res.mout$rd )
abline( h=res.mout$cutoff, lwd=2, col="red") 
abline( v=res.mout$cutoff, lwd=2, col="red")
llmout <- which( ( res.mout$md > res.mout$cutoff ) & (res.mout$rd > res.mout$cutoff) )
llmout
df$mout <- 0 
df$mout[ llmout ] <- 1 
df$mout <- factor( df$mout, labels = c("MvOut.No","MvOut.Yes"))
res.mout$cutoff
res.cat <- catdes(df[,c(2:8,10,18:19)],10)
res.cat$category
#The cars with Automatic transmission are overrepresented in multivariant outliers. And also there is a high percentage of automatic cars that are outliers (6.1%) in comparison to cars with other types of transmission. There is a relative low amount of semiautomatic cars that are outliers (20.81%) compared to the global amount of semiautomatic cars (38.45%).

summary(df[df$mout=="MvOut.Yes",])
summary(df)
#The cars that are outliers tend to be more expensive, have more mileage, have to pay more tax. The manufacturers Mercedes and VW have a low percentage of outliers cars.
```

# Profiling with FactoMineR

## Profiling of the numeric target variable "price"
```{r}
summary(df$price)
# The "variable to describe cannot have NA ###################################
res.condes<-condes(df,3, proba=0.01)

res.condes$quanti  # Global association to numeric variables
#The response variable has a strong correlation with the following variables: year, tax, mileage and mpg.
scatterplot(df$year,df$price)
scatterplot(df$tax,df$price)
scatterplot(df$mileage,df$price)
scatterplot(df$mpg,df$price)
res.condes$quali # Global association to factors
# P-values indicate whether the correlation is statistically different from 0 or not. p-values < 0.05 reject the null hypothesis (correlation statistically equal to 0). 
#The response variable has a strong correlation with the following variables: model, engineSize and transmission, according to the R2 stadistic value.

res.condes$category  # Partial association to significative levels in factors
#With this output we can see from different categories the mean difference in price compared to the mean price of the dataset
#The cars that have low mpg are more expensive.
#We can also see that the cars with an engine size = 4 has an estimate of +19400$
#We can also see that the cars with an engine size = 2.9 has an estimate of +19179$
#We can also see that the cars with an model=BMW- 7 Series has an estimate of +18054$
#We can also see that the cars with an model=Audi- Q8 has an estimate of +25943$
#We can also see that the cars with an engine size = 5.2 has an estimate of +32960$
#We can also see that the cars with an model=VW- Up has an estimate of -17472$
#We can also see that the cars with an engine size = 1.2 has an estimate of -15682$

```

## Profiling of the categorical target variable "Audi"

```{r}
summary(df$Audi)
# The "variable to describe cannot have NA
res.catdes<-catdes(df[,-c(1)],11, proba = 0.01)
#We exclude the model of the car from the analysis because it doesn't bring useful information.
res.catdes$quanti.var  # Global association to numeric variables
```
Miles per galon (mpg), price and mileage are statistically significant variables as they have a p-value less than 0.01. Despite that fact, the effect size associated with them is quite small as they have a small Eta2 value. This means that these variables are not quite significant at predicting if a car is an Audi or not.

```{r}
res.catdes$quanti # Partial association of numeric variables to levels of outcome factor
```
With this output we can see that Audi cars have a little more price and mileage than the global average and have fewer mpg than the global average. The opposite is true for cars that are not Audi.

```{r}
#mean(df$tax[which(df$Audi=="No")])-mean(df$tax[which(df$Audi=="Yes")])
res.catdes$test.chi2 # Global association to factors
res.catdes$category  # Partial association to significative levels in factors
```
With this final categorical analysis we can see that:
For cars that are not Audi:
    *We have smaller engine sizes overall. 
    *The percentage of cars with diesel and hybrid engines is slightly higher than the global mean.
    *We have more cars with a lower mileage.
    
For cars that are Audi:
    *The percentage of engines with a size of 1.4 is higher than the global mean (16.9  vs 6.9).
    *The percentage of Audis with a manual transmission is higher than the global mean (41 vs 35).



